---
title: "RWorksheet_Bagilidad#4c"
author: "Ace Bagilidad BSIT 2-C"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Use the dataset mpg.
a. Show your solutions on how to import a csv file into the environment.

```{r}

library(readr)

data_mpg <- read.csv("mpg.csv")
data_mpg 

```
b. Which variables from mpg dataset are categorical?

From the data mpg, the categorical data are the following:
1. Transmission(trans)        
2. Model                  
3. Displacement(displ)        
4. Drive Train(drv)                  
5. class

```{r}
data(mpg)
categorical_columns <- data_mpg %>% select_if(is.factor) %>% colnames()

# Print the categorical columns
print(categorical_columns)

```
c. Which are continuous variables?


2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
```{r}
# Assuming 'data_set' is already available in the environment
num_variables <- ncol(data)
print(paste("The number of variables in the data set is: ", num_variables))

```



