---
title: "RWorksheet_Bagilidad#4c"
author: "Ace Bagilidad BSIT 2-C"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Use the dataset mpg.
a. Show your solutions on how to import a csv file into the environment.

```{r}

library(readr)

data_mpg <- read.csv("mpg.csv")
data_mpg 

```
b. Which variables from mpg dataset are categorical?

From the data mpg, the categorical data are the following:

1. Transmission(trans)        
2. Model                  
3. Fuel type(fl)       
4. Drive Train(drv)                  
5. class
6. manufacturer

```{r}
library(dplyr)
data(mpg)
categorical_columns <- data_mpg %>% select_if(is.factor) %>% colnames()

# Print the categorical columns
print(categorical_columns)

```
c. Which are continuous variables?
1.X 
2.engine displacement(displ) 
3.year
4.cyl(number of cylinders)
5.city miles per gallon(cty)
6.highway miles per gallon(hwy)

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
Ans.
The manufacturer with the most model is:  dodge 
The model with the most variations is:  caravan 2wd
```{r}
manufacturer_table <- table(data_mpg$manufacturer)


#getting the most model
maxModel_Manufacturer <- as.character(names(manufacturer_table))[which.max(manufacturer_table)]

cat("The manufacturer with the most model is: ",maxModel_Manufacturer,"\n")

model_table <- table(data_mpg$model)
maxVariation_Model <- as.character(names(model_table))[which.max(model_table)]

cat("The model with the most variations is: ",maxVariation_Model)


```
2a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
library(dplyr)
models <- data.frame(Manufacturer = data_mpg$manufacturer,
                     Model = data_mpg$model)
models

unique_models <- unique(models)
unique_models



```



2b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
library(ggplot2)
factor_UniqueModels <- as.factor(unique_models$Manufacturer)
plotUnique_Models <- plot(as.factor(factor_UniqueModels),
                             main = "Unique Model per Manufacturer",
                             xlab = "Manufacturer", 
                             ylab = "Number of Unique Models",
                             cex.names = 0.6,
                             las = 2)

plotUnique_Models
```

2. Same dataset will be used. You are going to show the relationship of the model and the manufacturer.
2a.What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
ans. With the model on the x-axis of the chart and the manufacturer on the y-axis, it generates a scatterplot for the mpg dataset.
```{r}
library(ggplot2)

ggplot(data_mpg, 
       aes(model,manufacturer)) + 
   geom_point() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

```
2b. For you, is it useful? If not, how could you modify the data to make it more informative?
ans. Yes. It's very useful to know how many models each manufacturer produces.
To make it more interesting, I can put color codes on each point based on the classes variable.


3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results
```{r}
data_mpg_20 <- head(data_mpg,20)
data_mpg_20


ggplot(data = data_mpg_20, 
       aes(x= year, y=model,
       color = as.factor(model))) + geom_point(alpha = 0.7)+ labs(title = "Relationship between Model, Year, and Manufacturer",
       x = "Year",
       y = "Model",
       color = as.factor(data_mpg_20$model)) 
    

```
4.  Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result

```{r}
library(dplyr)

#group the number of cars per model

grouped_data <- data_mpg %>% group_by(model) %>% summarise(count = n()) 
grouped_data

```
4a. Plot using geom_bar() using the top 20 observations only. The graphs shoude have a title, labels and colors. Show code and results.
```{r}

observation_20 <- head(grouped_data,20)
observation_20

observation_plot <- ggplot(observation_20,
       aes(x = model,
           y = count,
           fill = model)) + 
  geom_bar(stat = "identity") +
   labs(title = "Top 20 Most Popular Car Models", 
       x = "Model",
       y = "Number of Cars") +
   theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

observation_plot

```
4b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}


observation_plot <- ggplot(observation_20,
       aes(x = reorder(model, count),
           y = count,
           fill = model)) + 
  geom_bar(stat = "identity") +
  coord_flip()+
  
  labs(title = "Number of Cars per Models", 
       x = "Model",
       y = "Number of Cars") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

observation_plot

```
5.Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.
```{r}
relationship_cyl_displ <- data.frame(Cylinder = data_mpg$cyl,
                                     Engine_Displacement = data_mpg$displ)
relationship_cyl_displ

relationship_plot <- ggplot(relationship_cyl_displ,
                            aes(x = Cylinder, y = Engine_Displacement, color = Engine_Displacement))+
                              geom_point()+
                              labs(title = "Relationship between number of cylinders and engine displacement",
                                   x = "No. of Cylinders",
                                   y = "Engine Displacement")+
  scale_color_gradient(low = "green",
high = "orange")+
                              theme(plot.title = element_text(hjust = 0.5))
relationship_plot
```
5a. How would you describe its relationship? Show the codes and its result.
Ans. When the number of cylinder increases the displacement also increases.

6.Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?

ans. Based on a color mapping to the city miles per gallon (cty), this output allows us to visually analyze 
the relationship between various city fuel economy levels and engine displacement (displ), 
as well as the highway miles per gallon (hwy), which helps us understand 
how fuel-efficient vehicles with various engine sizes are.
```{r}
relationship_displ_cyl <- ggplot(data_mpg,
                                 aes(x = displ, 
                                     y = hwy, 
                                     color = cty)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement",
       y = "Highway MPG") +
  scale_color_gradient(low = "green",
                       high = "orange")
  

relationship_displ_cyl

```

6. Import the traffic.csv onto your R environment.
```{r}

library(readr)
traffic_data <- read_csv("/cloud/project/RWorksheets_Bagilidad#4/RWorksheets#4C/traffic.csv")

```

6a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.

```{r}
observation_traffic <- colnames(traffic_data)

length_traffic <- length(observation_traffic)

cat("The number of observations are: ",length_traffic)
```
b. subset the traffic dataset into junctions. What is the R codes and its output?

```{r}
junction_sub1 <- subset(traffic_data, Junction == 1)
junction_sub1
junctions_sub2 <- subset(traffic_data, Junction == 2)
junctions_sub2
junctions_sub3 <- subset(traffic_data, Junction == 3)
junctions_sub3
junctions_sub4 <- subset(traffic_data, Junction == 4)
junctions_sub4

```

6c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
par(mfrow = c(1,4))
junctionPlot_1 <- ggplot(junction_sub1, 
                          aes(x = as.Date(junction_sub1$DateTime), 
                              y = Vehicles)) + 
   
  geom_line() +
  scale_x_date(date_labels = "%b-%Y") + theme(legend.position = "none") +
  labs(title = "Junction 1", x = "Time", y = "Number of Vehicles")

junctionPlot_1

junctionPlot_2 <- ggplot(junctions_sub2, 
                          aes(x = as.Date(junctions_sub2$DateTime), 
                              y = Vehicles)) + 
  geom_line() +
  scale_x_date(date_labels = "%b-%Y") + theme(legend.position = "none") +
  labs(title = "Junction 2", x = "Time", y = "Number of Vehicles")

junctionPlot_2

junctionPlot_3 <- ggplot(junctions_sub3, 
                          aes(x = as.Date(junctions_sub3$DateTime), 
                              y = Vehicles)) + 
  geom_line() +
  scale_x_date(date_labels = "%b-%Y") + theme(legend.position = "none") +
  labs(title = "Junction 3", x = "Time", y = "Number of Vehicles")

junctionPlot_3

junctionPlot_4 <- ggplot(junctions_sub4, 
                          aes(x = as.Date(junctions_sub4$DateTime), 
                              y = Vehicles)) + 
  geom_line() +
  scale_x_date(date_labels = "%b-%Y") + theme(legend.position = "none") +
  labs(title = "Junction 4", x = "Time", y = "Number of Vehicles")

junctionPlot_4



```
7. From alexa_file.xlsx, import it to your environment.

```{r}
library(readxl)
alexa_file_1 <- read_excel("/cloud/project/RWorksheets_Bagilidad#4/RWorksheets#4C/alexa_file (1).xlsx")

```

7a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.
```{r}
observation_alex <- nrow(alexa_file_1)


cat("Numbers of Observations: ", observation_alex,"\n")

columns_alex <- ncol(alexa_file_1)


cat("Numbers of Columns: ", columns_alex)

```


7b. group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.

```{r}
library(dplyr)
groupVar_alexa <- alexa_file_1 %>%
  group_by(variation) %>%
  summarise(Count_total = n())
groupVar_alexa
```

7c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.
```{r}
alexaPlot <- ggplot(alexa_file_1,
                    aes(x = variation,fill = variation))+
  geom_bar() +
  labs(title = "Variations",
       x = "Variation",
       y = "Count") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

alexaPlot
```
7d. d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.
```{r}
library(dplyr)

alexa_file_1$date <- as.Date(alexa_file_1$date)

alexa_file_1$month <- format(alexa_file_1$date, "%m")


alexa_fileMonth <-alexa_file_1%>%
  count(month)
alexa_fileMonth


alexa_file_monthly_reviews <- table(alexa_fileMonth)
alexa_file_monthly_reviews

alexa_file_plot <- ggplot(alexa_fileMonth, aes(x = month, y = n, group = 1)) +
  geom_line(color = "purple") +
  labs(title = "Number of Verified Reviews Over Time",
       x = "Date",
       y = "Number of Verified Reviews")

alexa_file_plot
```
e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a
graph to show its relationship. Show your solution and answer.
```{r}

alexa_variationRatings <- alexa_file_1 %>%
  group_by(variation) %>%
  summarise(average_rating = mean(rating))

alexa_variationRatings

topRatings <- alexa_variationRatings %>%
  filter(average_rating == max(average_rating))

topRatings 

# The walnut finish variation has the highest rating

ggplot(alexa_variationRatings, 
       aes(x = variation, 
       y = average_rating, 
       fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Ratings by Variation",
       x = "Variation",
       y = "Average Rating")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))


```